### 2019-12-31 docs/thesis-project/article.tex
Update article.tex

%%%%%%%%%%%%%%%%
% BEGIN HEADER %
%%%%%%%%%%%%%%%%
\title {Hives: File Survivability in P2P using Probabilistic Swarm Guidance
\thanks{This work has the support of the project MYRG2016-00097-FST of the University of Macau; by the Portuguese Fundação para a Ciência e a Tecnologia (FCT) through Institute for Systems and Robotics (ISR), under Laboratory for Robotics and Engineering Systems (LARSyS) project UID/EEA/50009/2019.}
}
\titlerunning{Hives}
\author{Francisco Barros, Daniel Silvestre \and Carlos Silvestre}
\authorrunning{F. Barros et al.}
\institute{Instituto Superior Técnico - Taguspark\newline Av. Prof. Doutor Cavaco Silva, 2744-016 Porto Salvo, Portugal
\email{fbarros@isr.ist.utl.pt, dsilvestre@isr.ist.utl.pt, csilvestre@umac.mo}\newline
\url{www.tecnico.ulisboa.pt}
}
\maketitle
%%%%%%%%%%%%%%
% END HEADER %
%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%
% BEGIN ABSTRACT %
%%%%%%%%%%%%%%%%%%
\begin{abstract}
Peer-to-Peer (P2P) systems have emerged as a potential choice to build large scale distributed storage systems, at a fraction of the cost of the alternative cloud approaches, which have better quality of service guarantees. We survey general-purpose P2P protocols, distributed file systems, cloud-assisted P2P protocols, and erasure-code algorithms. We then draft a solution for the problem of creating a bio-inspired P2P overlay to be used on a cloud-assisted P2P storage and backup system using probabilistic swarm guidance. We anticipate that our solution will have high network overhead, in order to provide self-healing properties to the files stored in the system.

@@

\subsection{Problem Description}
With the growth of the internet and the emergence of distributed systems, two large-scale computing paradigms have gained popularity due to their promise of virtually unlimited scalability. On the one hand, we have Peer-to-Peer (P2P) networking, which can be defined as a group of equally privileged peers, who contribute with a portion of their resources, to achieve common goals. These networks are popular among file sharing and streaming applications, due to their self-organized behavior, lack of centralization and, low-cost. On the other hand, Cloud platforms offer unmatched, on-demand, self-service, availability, and reliability, at a higher cost, and are currently trendy, with companies such as Amazon, Google, and Microsoft offering various ITaaS products to individuals and organizations alike. Cloud-based systems are centralized architectures, in which a large number of computers are clustered and managed by master entities, which possibly, become bottlenecks. Both paradigms can be used to deploy distributed file-storage or file-backup applications; despite the fact that P2P implementations are cheaper for both companies and their clients, these have a hard-time achieving 99.9\% availability seen in Cloud-based implementations, furthermore, depending on the P2P application there may also exist the risk of permanent file loss, making them somewhat unappealing for clients who want to store or backup their files remotely.

\subsection{Our approach}
The focus of this thesis is to create a structured P2P overlay to be used on a distributed backup system, where clients upload their files to remote peers, who are selling their storage space, without persistence guarantees. Probabilistic Swarm Guidance (PSG) and Markov Chain theories will be used, along with block-based file replication using erasure code algorithms and a reputation system, to hopefully create a network that is able to keep a file available at any given time, regardless of presence of attackers, network partitions and amount of churn, i.e., no matter how many peers leave/join an overlay.

\subsection
{Frequent approach} There has been much research on the topic of P2P overlays. There are two significant categories of overlays, Structured and Unstructured. In the former, the overlay has rigid rules regarding peer placement; they favor lookup speed but have higher maintenance overhead under churn, i.e., the process of peers leaving or joining the overlay. Some popular P2P Structured overlays include Chord, Tapestry, Kademlia, and P-Grid. In the latter, peer placement is random, which provides the overlay network with a higher degree of resilience and robustness in the advent of failures. Unstructured overlays disseminate information in a gossip-like fashion; thus, lookup operations are slower, resource management is less efficient, but overlay maintenance is often straightforward. Popular  Unstructured P2P overlays include FreeNet, Gnutella and BitTorrent. Research on multi-layer overlays has not yielded satisfactory results, and implementation is difficult \cite{sotart}, but research on bio-inspired overlays did have promising results, some of these include Self-Chord \cite{selfchord}. Finally, there have also been some attempts of transparently integrating Cloud Systems into P2P networks, e.g., CLOUDCAST \cite{cloudcast}, and vice-versa, e.g., Spotify and Wuala. To the best of our knowledge, there is no P2P, nor Cloud system using PSG as a way of routing data across the network nor as a way of ensuring file survivability.

\subsection{Proposal Goals}\label{subsec:intro}
This work seeks to address the problem of file survivability in P2P networks; our primary goal is to study, design, and evaluate a P2P overlay that uses probabilistic swarm guidance as a way to maintain a file available to service clients at all times. Steps will include:

\begin{itemize}
    \item Implement a Tracker server with light-weight responsibilities:
        \SubItem{Act as peer discovery service for clients;}
        \SubItem{Act as an indexer for clusters of peers (Hives);}
        \SubItem{Track peer reputations;}
    \item Implement an erasure-code algorithm used by Hives to replicate clients' files;
    \item Calculate the ideal steady-state distribution of file blocks, for a Hive, when:
        \SubItem{No history information on a peer is available;}
        \SubItem{Peer availability, reliability and other trust factors are known;}
    \item Hives converge to ideal block distribution in bounded-times using PSG;
    \item Efficiently update routing as Hives' topologies change over time;
    \item Balance the degree of each peer such that:
        \SubItem{Message exchange overhead is minimal;}
        \SubItem{Any changes to the probability of file loss are negligible;}
        \SubItem{Hive is resilient to massive node failures and resistant to churn;}
    \item Extra miles may include:
        \SubItem{Implement a trust system managed by the Tracker}
        \SubItem{Peers can reconstruct Hives without Tracker intervention;}
        \SubItem{Peers trust their neighbors independently of Tracker;}
        \SubItem{Hives can leverage dedicated Cloud-Storage if they become unreliable;}
\end{itemize}
%%%%%%%%%%%%%%%%%%%%
% END INTRODUCTION %
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
% BEGIN RELATED WORK %
%%%%%%%%%%%%%%%%%%%%%%
\newpage\section{Distributed File Systems}\label{sec:distributed-file-systems}
Distributed file system (DFS) enables individuals and organizations to store, access, and secure data on possibly remote locations, precisely as they would their local machines. It is common practice to replicate data at the file-level, block-level, or both, depending on the requirements of each system. For example, block-level replication may allow parallel execution of applications, such as is the case of Hadoop, Google, and Spark MapReduce algorithms. Here we present two DFS, which follow entirely different approaches when it comes to architecture and protocols. We focus only on the architectural implementation, their advantages or disadvantages. We do not consider parallel computation because our primary objective, in this thesis proposal, is the continuous availability of data in a system, to clients. Before we dive into them, it is essential to understand some desired characteristics of a DFS. In general, distributed systems of any kind should be \textbf{\textit{dependable}}, meaning that they should be able to avoid service failures that are more frequent and more severe than is acceptable. To be dependable is to be ready to provide the correct service (\textbf{\textit{availability}}), in a continuous manner (\textbf{\textit{reliability}}), with absence of catastrophic consequences to clients (\textbf{\textit{safety}}), e.g., losing a file, with no improper system alterations (\textbf{\textit{integrity}}) and, finally, to be able to undergo modifications and repairs at any time (\textbf{\textit{maintainability}}). Some of the features that contribute to the above characteristics are summarized below.

\textbf{\textit{Load-Balancing}} is the process of balancing communication, processing, and storage overhead among all nodes of the system. It helps to avoid bottlenecks, increasing operational performance, and helps to prevent catastrophic failures. Prevention of failures is also achieved through \textbf{\textit{fault tolerance}}, which enables a system to continue operating, possibly at a reduced level,  rather  than failing completely, when some part of it fails, to this end, some sort of \textbf{\textit{machine redundancy}} and \textbf{\textit{data redundancy}} may be important. As the amount of data passing in the network, being processed or stored in the system grows, or the number of clients using it increases, the capacity of a system to maintain its service without, or with little degradation, is called \textbf{\textit{scalability}}, it depends on the implementation of the system, and as a rule of thumb, centralized solutions are less scalable.

\subsection{Google File System}

Google File System (GFS) \cite{gfs} is designed to provide efficient and reliable access to data using clusters of commodity machines located in geographically spread data-centers. In particular, GFS aims to provide high data throughput, low latency, and reliability in the face of individual server failures. Because GFS runs over commodity hardware, the proposed model considers machine and network faults as being the norm rather than the exception.  This assumption is typical in modern distributed systems of any kind. GFS is the basis for an open-source project, which has seen continuous improvements over the years, called Hadoop Distributed File System (HDFS) \cite{hadoop}; we will not discuss it here because it holds many similarities to GFS with key differences being the approach to security management. One advantage of GFS over other architectures is that it leverages a distributed file system that tailors to the designers observation that, once written, files on google provided services, are only read, often sequentially, with random writes being practically non-existent, shifting optimization focus from latency to throughput, from random writes to appends when it comes to atomicity and no caching of data blocks. In the GFS architecture cluster consists of a single master and multiple chunk-servers. The Master maintains a mapping from files to chunks, the current location of chunks, and is responsible for failure detection over the chunk-servers as well as chunk-migration between them. Clients communicate, for reads and writes, with the Master for metadata operations and directly with the chunk-servers for data-bearing operations. Neither the client nor the chunk-servers cache file data because most of google's client applications stream through huge files. A read operation involves obtaining chunk handles and locations from the Master and then retrieving the chunks from one of the many available chunk-severs. For writes, the client splits the file into 64MB chunks and asks the Master to assign them an identifier, and then the client uploads his chunks to the chunk-servers indicated by the Master. See Fig.\ref{fig:gfs_arch} for a simplified architecture diagram.

GFS has proven its qualities in supporting large-scale data processing. However, small files have shown to lead to overloads on chunk-servers holding them because chunks have a fixed large value size of 64MB. The reasoning for GFS to use such substantial value is to reduce network overload, reduce the number of accesses to the Master, and reduce memory used at the Master due to mappings. Concerning our proposal, Hives, since convergence to a desired steady-state using Markov Chains theory requires a large number of chunks to exist in a P2P Hive, we might have to do file-level replication rather than the desired block-level when client files are too small.\newline

\begin{figure}[!ht]
\centering
\includegraphics[width=\textwidth]{images/gfs_paper.png}
\caption{GFS Architecture}
\label{fig:gfs_arch}
\end{figure}

\newpage\subsection{Ceph File System}
Ceph \cite{ceph} is an open-source, storage system that, like GFS, runs over commodity hardware. Ceph aims to deliver reliable, autonomous, distributed object, block, and file storage that is self-healing, self-managing, and has no single point of failure. The base architectural component of Ceph is the \textit{RADOS} object store. All content clients read and write to the system is represented as an object, irrespective of the object's original data type. \textit{RADOS} is a single logical entity composed of many \textit{Ceph Storage clusters}. Ceph architecture, Fig. \ref{fig:cfs_arch} is slightly more sophisticated than that of GFS, so we will summarize the components that are of interest to us. Inside a \textit{Ceph Storage cluster}, we find four types of logical machines, storage devices, abstracted with a layer called \textit{Object Software Daemon} (OSD), \textit{Monitors}, \textit{Managers}, and \textit{Metadata Servers}. \textit{Metadata Servers} are only required when a client his sharing his DFS with other clients, these servers contain information used for POSIX semantics, e.g., file metadata such as owner and last accessed timestamps. A small amount of \textit{Monitors} maintain a master copy of the storage cluster map with its current state. Monitors require high consistency and use Paxos to decide on the state of the cluster. \textit{Managers} maintains detailed information about placement groups, process metadata and host metadata. Finally, OSD store data on behalf of clients. Additionally, OSD utilize their nodes' resources to perform data replication, erasure coding, load-balancing, recovery, monitoring and reporting functions. Each Storage Cluster as anywhere from dozens to thousands of these OSDs.

The innovation Ceph brings is that the OSD within the same \textit{placement group}, within a cluster, are organized in a P2P fashion and are capable of autonomously carrying out replication and recovery tasks. \textit{Placement groups} are fragments of a logical \textit{pool}, and in turn, \textit{pools} are logical, dynamic partitions that define among other things, how much and which type of replication is done to objects that belong to it. When one OSD in a placement group suspects another as died or joined the cluster, it consults a \textit{Monitor} to obtain an up to date cluster map. Based on the cluster map and the \textit{CRUSH} algorithm, it is possible that this OSD will have to offload some of its content to other OSD. Clients also use \textit{CRUSH} to perform reads and writes to the system. To perform a write operation, for example, a client connects to a \textit{pool} in the storage cluster and runs \textit{CRUSH} with a cluster map retrieved from a Monitor. The result of the operation gives the client information on the placement group of the data he wishes to upload as well as the OSD he should contact to complete his request. When the target OSD receives the request, it runs \textit{CRUSH} to find the number of replicas that it should store. Then, it takes the object identifier, pool name, and the cluster map and uses \textit{CRUSH}, again, to discover the secondary OSDs to where he will replicate the write. The operation is completed when the primary OSD, receives acknowledgments from the remaining secondary OSD and acknowledges back to the client. \textit{CRUSH} is a light-weight pseudo-random placement algorithm that provides deterministic object quantization without lookups to central directories. Because it runs independently on clients and peers, it avoids the need to have the monitor constantly updating cluster to object mappings. Ceph is proof that distributed file systems can reliably be implemented using P2P behavior. Ceph OSDs perform a lot of self monitoring and management operations automatically and in a decentralized fashion, while being highly dependable. \cite{ceph_benchmarks}

\begin{figure}[!ht]
\centering
\includegraphics[width=0.8\textwidth]{images/cephfs.png}
\caption{CFS Architecture}
\label{fig:cfs_arch}
\end{figure}

\newpage\section{P2P Overlays}\label{sec:p2p-overlays}

In large-scale P2P systems, it is unfeasible and even undesirable for each peer in the network to know and explicitly interact with every other node connecting it to other peers. In this regard, a P2P overlay is a logical network that abstracts the physical network underneath it, typically IP-based, and where, single-hop edges represent links between pairs of peers, as depicted in Fig. \ref{fig:network_diagrams}. Before we survey related work in the field of P2P overlays, we first list some essential properties and technical jargon related to overlays and graph modeling, that helps us understand better the pros and cons of the research we are presenting in the next subsections of this thesis proposal.

@@

\subsection{Structured Overlays}
\subsubsection{Kademlia} \cite{kademlia} is structured overlay, yet it is fully decentralized, designed for key-value storage systems, which, according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology. Kademlia peer-identifiers and data-keys are built as in Chord\cite{chord}, using consistent hashing. Inserts and lookups on the overlay are unidirectional, according to the distance between two identifiers, calculated using XOR. The unidirectionality allows caching to be done along lookup paths, alleviating hot spots. Each peer in the overlay maintains a binary tree, routing table, whose leaves are lists of peers, called k-buckets, at distances $2^{x}$ to $2^{x+1}$, $x \in\hspace{1mm}\mathclose[0,160\mathclose]$ from itself, sorted according to a least recently seen policy. As a peer receives messages if the sender is a new acquaintance, that sender is placed on the proper k-bucket if there is enough space in it; otherwise, it will replace the least seen peer only if that peer fails to respond to a ping. By preferring old acquaintances, Kademlia seeks to enhance the stability of the overlay based on the observation that the longer a peer remains connected to the network, the higher the probability that it will remain connected at least another hour\cite{ssaroiu:msp2pfss}. This policy also prevents some denial of service attacks, where an enemy floods the network with new peers. The XOR metric gives Kademlia an advantage over other DHT-based implementations by making routing tables less rigid and by allowing peers in the network to learn useful routing information from discovery and lookup messages. Because of the existence of k-buckets, routing can also be done around failed nodes, providing some resistance to churn and partitioning.

@@

\newpage\subsection{Unstructured Overlays}
\subsubsection{Gia} \cite{gia} is a decentralized overlay protocol whose purpose is to overcome the drawbacks of Gnutella. Gnutella is the first-ever decentralized, unstructured P2P protocol. Gnutella's main issues are lack of scalability, as resource discovery bases itself on simple message flooding mechanisms and how easily some peers become overloaded when faced with high rates of aggregate queries, leading to system degradation. Gia proposes the use of super-nodes, which are now also part of the Gnutella protocol. The super-nodes receive and route queries to peers holding data, but these super-nodes and the construction of the topology around them is dynamic and adaptive. Gia recognizes the heterogeneity and resource constraints of the peers and also uses random walks to alter the topology adaptively as well as lookup operations. The results show that Gia is three to five orders of magnitude better than that of Gnutella. Random-walks by themselves, while minimizing the network overhead, are less likely to find appropriate responses to the performed lookup, unless they are biased to high-degree peers, but this can make peer overloading more frequent. Thus, to tackle both problems, Gia implements a topology adaptation that puts most peers within reach of high capacity peers, while simultaneously ensuring that they can handle the likely incoming requests. Lastly, all peers keep pointers to data items kept by their immediate one-hop neighbors.

@@

\subsection{CYCLON} CYCLON\cite{cyclon} is useable even in highly dynamic environments, providing peer degree symmetry, low diameter, low clustering, resilience to churn, and massive node failures. The focus of the authors was to provide a lightweight and simple protocol. It consists of having each peer keeping an ever-changing, partial view, with fixed-size \textit{c}, of the network in a structure called \textit{neighbors}. Periodically, independently and asynchronously, a peer contacts a random neighbor to exchange acquaintances. This process is called \textit{shuffling}. During a shuffle, a peer \textit{i}, first increments the age of all his known neighbors by one. Then, selects the oldest of all his neighbors, \textit{j}, and $l-1$ other random neighbors, forming \textit{s}. Then \textit{i} replaces \textit{j}'s network address by his own and sets his age to zero, in \textit{s}. After that, \textit{i} sends \textit{s} to \textit{j} and waits for \textit{j} to reply with \textit{s'}. Unlike \textit{s}, \textit{s'} is composed of a purely random subset of \textit{j}'s neighbors, without any age modifications. Upon reception of subsets \textit{s} and \textit{s'}, \textit{j} and \textit{i}, respectively, update their \textit{neighbors} by first using any empty slots in their partial views and then by replacing entries which they sent to the other, i.e., if \textit{i} has no more slots, he replaces, in its \textit{neighbors} structure, the entries that he sent in \textit{s} to \textit{j}, with the ones he received in \textit{s'} from \textit{j}. Note that, after \textit{i} initiates a shuffle with \textit{j}, \textit{i} becomes \textit{j}'s neighbor, but \textit{j} is no longer a neighbor to \textit{i}, i.e., the neighboring relation between \textit{i} and \textit{j} is reversed. Unlike many protocols, CYCLON links between peers are not bidirectional.

>>>>

%%%%%%%%%%%%%%%%
% BEGIN HEADER %
%%%%%%%%%%%%%%%%
\title {Hives: File Survivability in P2P using Probabilistic Swarm Guidance
\thanks{This work has the support of the project MYRG2016-00097-FST of the University of Macau; by the Portuguese Fundação para a Ciência e a Tecnologia (FCT) through Institute for Systems and Robotics (ISR), under Laboratory for Robotics and Engineering Systems (LARSyS) project UID/EEA/50009/2019.}
}
\titlerunning{Hives}
\author{Francisco Barros, Daniel Silvestre \and Carlos Silvestre}
\authorrunning{F. Barros et al.}
\institute{Instituto Superior Técnico - Taguspark\newline Av. Prof. Doutor Cavaco Silva, 2744-016 Porto Salvo, Portugal
\email{fbarros@isr.ist.utl.pt, dsilvestre@isr.ist.utl.pt, csilvestre@umac.mo}\newline
\url{www.tecnico.ulisboa.pt}
}
\maketitle
%%%%%%%%%%%%%%
% END HEADER %
%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%
% BEGIN ABSTRACT %
%%%%%%%%%%%%%%%%%%
\begin{abstract}
Peer-to-Peer (P2P) systems have emerged as a potential choice to build large scale distributed storage systems, at a fraction of the cost of the alternative cloud approaches, which have better quality of service guarantees. We survey  distributed file systems, P2P overlays, cloud-assisted P2P protocols, and erasure-code algorithms. We then draft a solution for the problem of creating a P2P overlay to be used on a cloud-assisted P2P backup system that uses probabilistic swarm guidance and aims to ensure file durability. We anticipate that our solution will have high network overhead, in order to provide good self-healing properties to the files stored in the system.

@@

\subsection{Problem Description}
With the growth of the internet and the emergence of distributed systems, two large-scale computing paradigms have gained popularity due to their promise of virtually unlimited scalability. On the one hand, we have Peer-to-Peer (P2P) networking, which can be defined as a group of equally privileged peers, who contribute with a portion of their resources, to achieve common goals. These networks are popular among file sharing and streaming applications due to their self-organized behavior, lack of centralization and, low-cost. On the other hand, Cloud platforms offer unmatched, on-demand, self-service, availability, and reliability, at a higher cost, and are currently trendy, with companies such as Amazon, Google, and Microsoft offering various ITaaS products to individuals and organizations alike. Cloud-based systems are centralized architectures, in which a large number of computers are clustered and managed by master entities, which possibly, become bottlenecks. Both paradigms can be used to deploy distributed file-storage or file-backup applications; although P2P implementations are cheaper for both companies and their clients, these have a hard-time achieving availability levels seen in Cloud implementations. P2P systems also have a higher inherent risk of permanent file loss, making them somewhat unappealing for clients.

\subsection{Frequent approach}
P2P research often focuses on optimizing the overlay topology to provide load-balancing, read/write efficiency, and partition avoidance, implicitly assuming that reliable overlays provide reliable systems, in the case of distributed storage systems, replication or erasure coding is usually written on top of the P2P protocol to ensure durability. The most common approach for distributed storage systems, however, is done on the cloud using highly available machines, with expensive file or block-level replication, continuously monitored by failure detection entities to ensure the replication level is always above some threshold; While effective, the latest architecture is very costly and inefficient. Finally, there have also been some attempts of transparently integrating Cloud Systems into P2P networks, e.g., Cloudcast \cite{cloudcast}, and vice-versa, e.g., Spotify\cite{spotify} and Wuala\cite{wuala}. To the best of our knowledge, there is no system using PSG as a way of simultaneously deciding data routes, failure-detection, load-balancing, and, thus, file survivability in distributed storage systems.

\subsection{Our approach}
The focus of this thesis is to create a distributed backup system. Probabilistic Swarm Guidance (PSG) and Markov Chain theories will be used, along with erasure coding. We hope not to lose the files regardless of the amount of churn\footnote{Property of P2P systems, which regards peers joining or leaving a P2P network} or other experienced faults. Furthermore, to boost the availability and reliability files, cloud servers might be sporadically used to assist the P2P networks when these are facing critical conditions. The main difference between our approach and the frequent ones is that we use a layered solution, where a typical client-server approach facilitates lookup operations and provides system-wide but file durability, load-balancing, and failure detection is the responsibility of multiple P2P overlays specifically tailored for backing up files.

\subsection{Proposal Goals}\label{subsec:intro}
This proposal seeks to address the problem of file survivability in P2P networks; our primary goal is to design and evaluate a P2P overlay that uses probabilistic swarm guidance and erasure-coding as a mean to provide file survivability. Our secondary goal is to integrate instances of this overlay on a broader centralized system that provides fast look-up operations and increases availability and reliability by making connections to the cloud when needed. Steps include:

\begin{itemize}
    \item Identify a steady-state, $Q$, providing optimal robustness and load-balancing
    \item Identify a transition matrix, $M$, that allows peers to achieve $Q$ collectively
    \item Optimize $M$ with respect to peer connectivity degree
    \item Optimize updates to $M$ as overlay topology changes over time
    \item Implement a replication model that exploits the properties of the $M$
    \item Evaluate solution performance in a Simulation or on in real life scenario
\end{itemize}
%%%%%%%%%%%%%%%%%%%%
% END INTRODUCTION %
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
% BEGIN RELATED WORK %
%%%%%%%%%%%%%%%%%%%%%%
\newpage\section{Distributed File Systems}\label{sec:distributed-file-systems}
Distributed file system (DFS) enables individuals and organizations to store, access, and secure data on possibly remote locations, as they would their local machines. It is common practice to do file-level or block-level, but it is rare to use erasure-coding. The choice depends on the solution requirements and the broadness of the system. For example, block replication allows parallel execution of applications using MapReduce algorithms, and because of that, Google File System\cite{gfs} (GFS) uses block-level replication. We present two different implementations running on commodity hardware and focus on their architecture, advantages, or disadvantages. Before that, it is essential to understand some desired characteristics of a DFS. In general, distributed systems of any kind should be \textbf{\textit{dependable}}, meaning that they should be able to avoid service failures that are more frequent and more severe than is acceptable. To be dependable is to be ready to provide the correct service (\textbf{\textit{availability}}), in a continuous manner (\textbf{\textit{reliability}}), with absence of catastrophic consequences to clients (\textbf{\textit{safety}}), e.g., losing a file, with no improper system alterations (\textbf{\textit{integrity}}) and, finally, to be able to undergo modifications and repairs at any time (\textbf{\textit{maintainability}}). Some of the features that contribute to the above characteristics are the following: \textbf{\textit{load-Balancing}} is the process of balancing communication, processing, and storage overhead among all nodes of the system. It helps to avoid bottlenecks, increasing operational performance, and helps to prevent catastrophic failures. Prevention of failures is also achieved through \textbf{\textit{fault tolerance}}, which enables a system to continue operating, possibly at a reduced level,  rather  than failing completely, when some part of it fails, to this end, some sort of \textbf{\textit{machine redundancy}} and \textbf{\textit{data redundancy}} may be important. As the amount of data passing in the network, being processed or stored in the system grows, or the number of clients using it increases, the capacity of a system to maintain its service without, or with little degradation, is called \textbf{\textit{scalability}}, it depends on the implementation of the system, and as a rule of thumb, centralized solutions are less scalable.

\subsection{Google File System}
GFS wants to provide efficient and reliable access to data using clusters located in geographically spread data-centers. In particular, GFS aims to provide high data throughput, low latency, and reliability in the face of individual server failures. Because GFS model considers machine and network faults as being the norm, as is, nowadays typical. GFS serves as inspiration for a massively adopted open-source project called Hadoop Distributed File System\cite{hadoop}; we will not discuss it here because it holds many similarities to GFS with key differences being the approach to security management. Based on the author's observation that once written, files on google services, are only read, often sequentially, with random writes being practically non-existent, GFS chooses to optimize throughput over latency, appends over random writes and deliberately skips data block caching. GFS clusters consists of a single master and multiple chunk-servers. The Master maintains a mapping from files to chunks, the current location of chunks, and is responsible for failure detection over the chunk-servers as well as chunk-migration between them. Clients communicate, for reads and writes, with the Master for metadata operations and directly with the chunk-servers for data-bearing operations. A read operation involves obtaining chunk handles and locations from the Master and then retrieving the chunks from one of the many available chunk-severs. For writes, the client splits the file into 64MB chunks and asks the Master to assign them an identifier, and then the client uploads his chunks to the chunk-servers indicated by the Master. GFS has proven its qualities in supporting large-scale data processing. However, small files have shown to lead to overloads on chunk-servers holding them because chunks have a fixed large value size of 64MB. The reasoning such substantial size value is to reduce network overhead, reduce the number of accesses to the Master, and reduce memory used at the Master due to mappings. Concerning our proposal, since convergence to a desired steady-state is likely better achieved when more parts exist on the P2P network, we might consider using file replication rather than erasure-coding when files are small.

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth]{images/gfs_paper.png}
\caption{GFS Architecture}
\label{fig:gfs_arch}
\end{figure}

\subsection{Ceph File System}

\begin{figure}[t]
\centering
\includegraphics[width=0.6\textwidth]{images/cephfs.png}
\caption{CFS Architecture}
\label{fig:cfs_arch}
\end{figure}

Ceph \cite{ceph} is an open-source system that aims to deliver reliable, autonomous, distributed object, block, and file storage that is self-healing, self-managing, and has no single point of failure. The base architectural component of Ceph is the RADOS object store. All clients' reads and writes to the system are represented as objects, irrespective of the objects' original data types. RADOS is a single logical entity composed of many storage clusters. Ceph architecture is sophisticated, so, we summarize the main components. Inside a storage cluster, we find four types of logical machines, storage devices, abstracted with a layer called \textit{object software daemon} (OSD), \textit{monitors}, \textit{managers}, and \textit{metadata servers}. Metadata servers contain information used for POSIX semantics, e.g., file owner or last accessed timestamps. A small number of monitors maintain a master copy of the storage cluster map with its current state. Monitors use Paxos to decide on the state of the cluster. Managers maintain detailed information about placement groups, process and host metadata. Finally, OSD store data on behalf of clients. Additionally, OSD utilize their nodes' resources to perform replication, erasure coding, load-balancing, recovery, monitoring, and reporting functions. Each Storage Cluster as anywhere from dozens to thousands of these OSDs. Ceph's innovation is that OSD within the same \textit{placement group}, within a cluster, are organized in a P2P fashion and are capable of autonomously carrying out recovery tasks. Placement groups are fragments of a logical \textit{pool}, and in turn, pools are logical, dynamic partitions that define among other things, how much and which type of replication is done to objects that belong to it. When one OSD in a placement group suspects another as left or joined the cluster, it consults a Monitor to obtain an up to date cluster map. After running the CRUSH algorithm\cite{crush} over the cluster map, it is possible that the OSD will offload some of its content to another OSD. Clients also use CRUSH to perform reads and writes to the system. To perform a write operation, for example, a client connects to a pool in the storage cluster and runs CRUSH with a cluster map retrieved from a Monitor. The result of the operation gives the client information on the placement group of the data he wishes to upload as well as the OSD he should contact to complete his request. When the target OSD receives the request, it takes the object identifier, pool name, and the cluster map and uses CRUSH, to discover the number of replicas that it should store and to which secondary OSDs he must replicate the write. The operation is completed when the primary OSD, receives acknowledgments from the remaining secondary OSD and acknowledges back to the client. Because operations runs independently on clients and OSDs, there is no need to have the monitor constantly updating cluster to object mappings. Ceph is proof that DFS can reliably be implemented using P2P behavior. Ceph OSDs perform a lot of self-monitoring and management operations and in a decentralized fashion while being highly dependable\cite{ceph_benchmarks}; we hope to be able to replicate these achievements in our approach.

\section{P2P Overlays}\label{sec:p2p-overlays}
In large-scale P2P systems, it is unfeasible and even undesirable for each peer in the network to know and explicitly interact with every other node connecting it to other peers. In this regard, a P2P overlay is a logical network that abstracts the physical network underneath it, typically IP-based, and where, single-hop edges represent links between pairs of peers, as depicted in Fig. \ref{fig:network_diagrams}. Before we survey related work in the field of P2P overlays, we first list some essential properties and technical jargon related to overlays and graph modeling, that helps us understand better the pros and cons of the research we are presenting in the next subsections of this thesis proposal.

@@

\subsection{Structured Overlays}
Structured overlay ensures peer placement follows rigid rules in order to speed up lookup operations, commonly using Distributed Hash Tables or Tree implementations. However, for this reason, these overlays also have higher maintenance costs due to churn, making them less ideal for highly dynamic environments.

\subsubsection{Kademlia} \cite{kademlia} is fully decentralized and designed for key-value storage systems, which, according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology. Kademlia peer-identifiers and data-keys are built as in Chord\cite{chord}, using consistent hashing. Inserts and lookups on the overlay are unidirectional, according to the distance between two identifiers, calculated using XOR. The unidirectionality allows caching to be done along lookup paths, alleviating hot spots. Each peer in the overlay maintains a binary tree, routing table, whose leaves are lists of peers, called k-buckets, at distances $2^{x}$ to $2^{x+1}$, $x \in\hspace{1mm}\mathclose[0,160\mathclose]$ from itself, sorted according to a least recently seen policy. As a peer receives messages if the sender is a new acquaintance, that sender is placed on the proper k-bucket if there is enough space in it; otherwise, it will replace the least seen peer only if that peer fails to respond to a ping. By preferring old acquaintances, Kademlia seeks to enhance the stability of the overlay based on the observation that the longer a peer remains connected to the network, the higher the probability that it will remain connected at least another hour\cite{ssaroiu:msp2pfss}. This policy also prevents some denial of service attacks, where an enemy floods the network with new peers. The XOR metric gives Kademlia an advantage over other DHT-based implementations by making routing tables less rigid and by allowing peers in the network to learn useful routing information from discovery and lookup messages. Because of the existence of k-buckets, routing can also be done around failed nodes, providing some resistance to churn and partitioning.

@@

\subsection{Unstructured Overlays}
Unstructured overlays usually disseminate information in a gossip-like fashion; thus, lookup operations are slower than structured overlays, resource management is less efficient, but overlay maintenance is often straightforward. Because peer placement is arbitrary, they tend to provide better scalability and a higher degree of resilience in the advent of failures.

\subsubsection{Gia} \cite{gia} is a decentralized overlay protocol whose purpose is to overcome the drawbacks of Gnutella. Gnutella is the first-ever decentralized, unstructured P2P protocol. Gnutella's main issues are lack of scalability, as resource discovery bases itself on simple message flooding mechanisms and how easily some peers become overloaded when faced with high rates of aggregate queries, leading to system degradation. Gia proposes the use of super-nodes, which are now also part of the Gnutella protocol. The super-nodes receive and route queries to peers holding data, but these super-nodes and the construction of the topology around them is dynamic and adaptive. Gia recognizes the heterogeneity and resource constraints of the peers and also uses random walks to alter the topology adaptively as well as lookup operations. The results show that Gia is three to five orders of magnitude better than that of Gnutella. Random-walks by themselves, while minimizing the network overhead, are less likely to find appropriate responses to the performed lookup, unless they are biased to high-degree peers, but this can make peer overloading more frequent. Thus, to tackle both problems, Gia implements a topology adaptation that puts most peers within reach of high capacity peers, while simultaneously ensuring that they can handle the likely incoming requests. Lastly, all peers keep pointers to data items kept by their immediate one-hop neighbors.

@@

\subsection{Cyclon} Cyclon\cite{cyclon} is useable even in highly dynamic environments, providing peer degree symmetry, low diameter, low clustering, resilience to churn, and massive node failures. The focus of the authors was to provide a lightweight and simple protocol. It consists of having each peer keeping an ever-changing, partial view, with fixed-size \textit{c}, of the network in a structure called \textit{neighbors}. Periodically, independently and asynchronously, a peer contacts a random neighbor to exchange acquaintances. This process is called \textit{shuffling}. During a shuffle, a peer \textit{i}, first increments the age of all his known neighbors by one. Then, selects the oldest of all his neighbors, \textit{j}, and $l-1$ other random neighbors, forming \textit{s}. Then \textit{i} replaces \textit{j}'s network address by his own and sets his age to zero, in \textit{s}. After that, \textit{i} sends \textit{s} to \textit{j} and waits for \textit{j} to reply with \textit{s'}. Unlike \textit{s}, \textit{s'} is composed of a purely random subset of \textit{j}'s neighbors, without any age modifications. Upon reception of subsets \textit{s} and \textit{s'}, \textit{j} and \textit{i}, respectively, update their \textit{neighbors} by first using any empty slots in their partial views and then by replacing entries which they sent to the other, i.e., if \textit{i} has no more slots, he replaces, in its \textit{neighbors} structure, the entries that he sent in \textit{s} to \textit{j}, with the ones he received in \textit{s'} from \textit{j}. Note that, after \textit{i} initiates a shuffle with \textit{j}, \textit{i} becomes \textit{j}'s neighbor, but \textit{j} is no longer a neighbor to \textit{i}, i.e., the neighboring relation between \textit{i} and \textit{j} is reversed. Unlike many protocols, Cyclon links between peers are not bidirectional.
